Characterizing Stellar Streams by ML

### 2023-08-23 src/tex/ms.tex
[showyourwork] overleaf sync

% Change the section names
\renewcommand{\sectionautorefname}{\S\!}%
\renewcommand{\subsectionautorefname}{\S\!\!}%
\renewcommand{\subsubsectionautorefname}{\S\!}%
\renewcommand{\equationautorefname}{Eq.}

@@

    % subsection framing_the_data (end)

@@

    % subsection likelihood_setup (end)

@@

        \subsubsection{Linking to Astrometrics: the Distance Modulus and the Parallax}

@@

@@

    % subsection photmetric background (end)

@@

        % subsection subsubsection parameter_bounds (end)

@@

    \subsection{Model Implementation, aka the code} \label{sub:model_implementation}

        In prior sections we

        Thus far we have described the mathematical model.
        Now we describe

@@

        \begin{enumerate}
            \item A network per model component, e.g. one for the weights, one for the stream track and width, one for the background distribution.
            \item MLP for the components - lin / tanh + dropout
            \item Normalizing flow - specific pieces
            \item connective priors
            \item 
            \item Architecture in the Appendix
        \end{enumerate}

        \begin{table}
            \centering
            \caption{MLP structure}
            \begin{tabular}{@{}rrcl@{}}
            \toprule
            & Layer & Shape & Param \# \\
            \midrule
            & Linear & (None, 1, F) & 1  \\
            & Tanh & (None, 1, F) & 1  \\
            & Dropout & (None, 1, F) & 0  \\
            & Linear & (None, 1, F) & 1  \\
            \bottomrule
            \end{tabular}
        \end{table}

@@

    % subsection test_on_a_synthetic_stream (end)

@@

% subsection gd1 (end)

@@

        % subsection streamfinder (end)

@@

        % subsection via_machinae (end)

@@

        % subsection uniform_modeling_of_13_stellar_streams (end)

>>>>

% Change the section names
\renewcommand{\sectionautorefname}{\S\!}%
\renewcommand{\subsectionautorefname}{\S\!\!}%
\renewcommand{\subsubsectionautorefname}{\S\!}%
\renewcommand{\equationautorefname}{Eq.}
\renewcommand{\tableautorefname}{Tab.}

@@

    % sub: framing_the_data (end)

@@

    % sub: likelihood_setup (end)

@@

        \subsubsection{Linking to Astrometrics: the Distance Modulus and the Parallax} \label{ssub:linking_to_astrometrics}

@@

        % ssub: linking_to_astrometrics (end)

@@

@@

        % ssub: parameter_bounds (end)

@@

        % ssub: region_prior (end)

    % sub: priors (end)

    \vspace{5pt}
    \subsection{Model Implementation, aka the code} \label{sub:model_implementation}

@@

        In prior sections we discussed the mathematical implementation of each component of the Mixture Density Network. Here we make brief remarks on the code implementation and architecture of the neural networks.
        % Rephrase: I don't like this topic sentence.

        The neural networks are built with the \package{PyTorch} stack \TODO{[cite]}. We wrap the network \texttt{pytorch.Module} in a custom framework, allowing the Bayesian likelihood, prior, evidence, etc. to be bundled with the network's forward step. This framework allows for a high degree of modularity. Component models, e.g. the stream model, of the mixture are easily added to (and moved from) as items in the larger mixture model. Probability methods understand how to traverse the mixture model, calling and combining the appropriate methods of the constituent models, so it unnecessary to write custom loss functions. 
        Another benefit of the modular framework is the ease of building and experimenting with different mixture models. Furthermore, each model component may be saved and serialized separately from the other components, allowing for
        portability of the components between models. For example, if a stream has a prominent extra-tidal feature, e.g. the spur of \stream{GD-1}, a fiducial model of just the thin stream may be trained, then the stream (and background) model components may be saved and used as the initialization to the stream (and background) of a more complete model that includes the spur. This modularity also has a cost: by allowing each model component (including the categorical mixture model itself) to contain its own network we decrease the cross-talk between nodes in separated networks. The networks communicate only through the backwards propagation of the loss function, increasing the required training time of the models. We note that the code framework, being built on \package{PyTorch}, easily permits a single model with single network. Having found a good model by experimenting with the modular layout, the model may be reproduced and re-trained as a single network. Alternatively, though we leave this to a future version of the code, it is possible to ``fuse" the mixture, providing one network whose outputs are apportioned appropriately to each constituent model.

        The models permit user-defined priors to be passed at model instantiation.
        Priors specific to a single model are contained on that model, while priors connecting two (or more) models are held on the mixture model object. As an example of the latter, see \autoref{ssub:linking_to_astrometrics}, which connects the distance component of the astrometric and photometric stream models.

        The custom model framework combines for each model a neural network with Bayesian probability functions. The latter are set by the model type, e.g. the Gaussian model has a Gaussian likelihood function, but the former, the neural network, is set and determined by the user. So long as the network has the correct number of inputs (e.g. one for $\phi_1$) and outputs (one per parameter), any feed-forward network may be used. Our default choice, and for which we provide a helper function, is an MLP with a sequence of (\texttt{Linear} \textrightarrow \texttt{Tanh} \textrightarrow \texttt{Dropout}) blocks. This is illustrated in \autoref{table:mlp_structure}.
        % FIXME! reference not being picked up correctly.
        One model for which the lin-tanh network cannot be used is the normalizing flow
        used in the photometric background model (\autoref{sub:photometric_model_off_stream}). Here we use the \package{nflows} package, with a conditional diagonal normal base distribution and composite transform network of a reverse permutation and masked affine auto-regressive transform layers. As noted in our data availability statement, all the codes used in this paper are bundled by \href{https://github.com/showyourwork/showyourwork}{\package{showyourwork}} in a public repository at $\GitHubURL{}$. We will describe for each stream model specifics relevant to that stream but refer readers to the publicly-available code itself for code implementation details.
        % TODO: Maybe more details in the Appendix?

        \begin{table}
            \centering
            \caption{\todo[inline]{MLP structure}}
            \label{table:mlp_structure}
            \begin{tabular}{@{}rrcl@{}}
            \toprule
            & Layer & Shape & Param \# \\
            \midrule
            & Linear & (None, 1, H) & 1  \\
            & Tanh & (None, H, H) & 1  \\
            & Dropout & (None, H, H) & 0  \\
            & Linear & (None, H, H) & 1  \\
            & Tanh & (None, H, H) & 1  \\
            & Dropout & (None, H, H) & 0  \\
            & Linear & (None, H, F) & 1  \\
            & Tanh & (None, F, F) & 1  \\
            & Dropout & (None, F, F) & 0  \\
            \bottomrule
            \end{tabular}
        \end{table}

@@

    % sub: test_on_a_synthetic_stream (end)

@@

% sub: gd1 (end)

@@

        % sub: streamfinder (end)

@@

        % sub: via_machinae (end)

@@

        % sub: uniform_modeling_of_13_stellar_streams (end)
